### 2018-04-13 paper/evaluation.tex
a pass for the paper

% describe benchmarks:
We evaluated \sys and \duality on 4,309 CHC systems generated from
programs in the SV-COMP 2015~\cite{svcomp15} verification benchmark
suite.
%
To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
verification framework with its default settings, set to timeout at 90
seconds.
%
We used the benchmarks in SV-COMP 2015~\cite{svcomp15} because they
were used to evaluate \duality in previous work~\cite{mcmillan14}.

>>>>

% describe benchmarks:
We evaluated \sys and \duality on 4,309 CHC systems generated from
programs in the SV-COMP 2015~\cite{svcomp15} verification benchmark
suite.
%
To generate CHC systems, we ran the \seahorn~\cite{gurfinkel15}
verification framework with its default settings which not inlining procedure call and each 
loop-free fragment is a clause, set to timeout at 90
seconds.
%
We used the benchmarks in SV-COMP 2015~\cite{svcomp15} because they
were used to evaluate \duality in previous work~\cite{mcmillan14}.

### 2018-04-13 paper/introduction.tex
a pass for the paper

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that we refer to as
\emph{Clause-Dependence Disjoint} (CDD).
%
The class consists of all recursion-free CHC systems such that each
clause in the system has at most one dependency on the same relational
predicate.
%
The key characteristic of this class is that when an arbitrary
recursion-free system is reduced to a CDD system and to a system from
a different directly solvable class, the CDD system is frequently the
smaller of the two.
%
Therefore, solving recursion-free systems by reducing them to CDD form
is often less computationally expensive than reducing them to a
system in a different class.

>>>>

% contribution of this paper: CDD systems:
The first contribution of this paper is the introduction of a novel
class of directly solvable systems that we refer to as
\emph{Clause-Dependence Disjoint} (CDD).
%
The formal definition of CDD is given at ~\autoref{defn:cdds}.
%
CDD is a strict superset of the union of all previous introduced directly
solvable systems.
%
The key characteristic of this class is that when an arbitrary
recursion-free system is reduced to a CDD system and to a system from
a different directly solvable class, the CDD system is frequently the
smaller of the two.
%
Therefore, solving recursion-free systems by reducing them to CDD form
is often less computationally expensive than reducing them to a
system in a different class.

### 2018-04-13 paper/overview.tex
a pass for the paper

%first introduce the question
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$ (\autoref{fig:chc}), that models the safety of a small program,
\cc{dblAbs} (\autoref{fig:multicall-code}).
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a CDD system and
how \sys can solve it by encoding it into binary interpolants.
%
In \autoref{sec:not-in}, we illustrate that $\mcchc$ is not in
directly solvable classes introduced in previous work.
%

@@

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.3\textwidth] {%
      \caption{\cc{dblAbs}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    {\lstinputlisting[
      basicstyle=\small,
      xleftmargin=\parindent,
      numbers=left,
      morekeywords={def, return, if, else, assert}]{code/dblAbs.java}
        % \input{code/dblAbs.java}}
    }
    %
    \ffigbox[.66\textwidth]{%
      \caption{A CHC system that models the safety condition of $\cc{dblAbs}$,
    named $\mcchc$.}}{%
      \label{fig:chc}
      \begin{align}
        % semantic constraint of dbl:
        \cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
        % semantic constraint of line 3:
        \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
        % semantic constraint of line 5 (then branch)
        \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
        % semantic constraint of line 6 (else branch)
        \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
        % assign branch
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
        % semantic constraint of main procedure:
        \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
        \cc{dbl}(\cc{x}, \cc{d})
        \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
        %
        False &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
      \end{align}
    }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{dblAbs} is a small procedure that doubles the absolute value of
its input and stores the result in \cc{res}.
% state the safety problem:
The program also asserts that \cc{res} is greater than or equal to $0$
before exiting.
% talk about solving Horn Clauses:
Verifying this assertion reduces to solving a recursion-free
CHC system over a set of uninterpreted predicates that represent
the control locations in \cc{dblAbs}.
%
In particular, one such system, $\mcchc$, is shown in
\autoref{fig:chc}.
%
%
While $\mcchc$ has been presented as the result
of a translation from \cc{dblAbs}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

@@

For example, consider $\cc{L}_9$. By the time \sys attempts to
synthesize an interpretation for $\cc{L}_9$ it will have solutions for
$\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
%
Possible interpretations of these predicates are shown in
\autoref{fig:ex-graph}.
%
The pre-formula is constructed from the bodies of clauses where
$\cc{L}_9$ is the head.
%
Each relational predicate, $P$,
is replaced by a corresponding boolean indicator variable, $\cc{b}_P$.
%
Each boolean indicator variable implies the solution for its
predicate,
encoded as the disjunction of the negation of the boolean
indicator variable and the solution.
%
In particular, the pre-formula for $\cc{L}_9$ is constructed from
clauses (5) and (6):
\begin{gather}
  ((\cc{b}_{\cc{L}_6} \land \cc{abs'} = \cc{n})
  \lor
  (\cc{b}_{\cc{L}_8} \land \cc{abs'} = -\cc{n}))
  \land
  (\neg \cc{b}_{\cc{L}_6} \lor \cc{n} \ge 0)
  \land
  (\neg \cc{b}_{\cc{L}_8} \lor \cc{n} < 0)
\end{gather}
%
The post-formula is constructed from clauses that transitively depend
on $\cc{L}_9$. Again, we replace relational predicates by
corresponding boolean indicators. However, we omit the boolean indicator for
$\cc{L}_9$. The post-formula is composed from clauses (1), (7), and
(8):
\begin{gather}
  (\neg \cc{b}_{\cc{dbl}} \lor \cc{d} = 2*\cc{x})
  \land
  (\neg \cc{b}_{\cc{main}} \lor
    (\cc{b}_{\cc{dbl}}
    \land \cc{abs'}= \cc{x}
    \land \cc{res}=\cc{d} ))
  \land
  (\cc{b}_{\cc{main}} \land \cc{res'} < 0)
\end{gather}
%
Interpolating the pre and post formulas yields an interpretation
of $\cc{L}_9$: $\cc{abs'} \geq 0$.
%
The procedure for solving a CDD system is described in formal detail
in \autoref{sec:solve-cdd}.

@@

%% talk about expansion:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.45\textwidth] {%
      {\caption{$\mcchc$ as a directed hypergraph. Each relational
          predicate is depicted as a graph node while each clause is
          represented by a hyperedge. Each hyperedge is labeled by the
          fact in the corresponding CHC\@. Each node has a valid
          corresponding interpretation, written in braces.
      }\label{fig:ex-graph}}}
      {\input{fig/example}}
    \ffigbox[.45\textwidth]
      {\caption{The hierarchy of classes of recursion-free CHC
          systems. Body Disjoint and Linear systems are subsumed by
          CDD systems. Solving Directly Solvable CHC systems is in
          co-NP while solving general, recursion-free systems is
          co-NEXPTIME Complete.
      }\label{fig:hierarchy}}
      {\input{fig/hierarchy}}
  \end{floatrow}
\end{figure}

>>>>

%first introduce the question
In \autoref{sec:running-ex}, we describe a recursion-free CHC system,
$\mcchc$ (\autoref{fig:chc}), that models the safety of program \cc{dblAbs} (\autoref{fig:multicall-code}).
%
In \autoref{sec:solve-ex}, we show that $\mcchc$ is a CDD system and
how \sys can solve it by encoding it into binary interpolants.
%
In \autoref{sec:not-in}, we illustrate that $\mcchc$ is not in
directly solvable classes introduced in previous work.
%

@@

% include code of running example and some figure:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.3\textwidth] {%
      \caption{\cc{dblAbs}: an example hierarchical program.} %
      \label{fig:multicall-code} }
    {\lstinputlisting[
      basicstyle=\small,
      xleftmargin=\parindent,
      numbers=left,
      morekeywords={def, return, if, else, assert}]{code/dblAbs.java}
        % \input{code/dblAbs.java}}
    }
    %
    \ffigbox[.66\textwidth]{%
      \caption{A CHC system that models the safety condition of $\cc{dblAbs}$,
    named $\mcchc$.}}{%
      \label{fig:chc}
      \begin{align}
        % semantic constraint of dbl:
        \cc{dbl}(\cc{x}, \cc{d}) &\gets \cc{d} = 2 * \cc{x} \\
        % semantic constraint of line 3:
        \cc{L}_4(\cc{n}, \cc{abs}) &\gets  \cc{abs} = 0 \\
        % semantic constraint of line 5 (then branch)
        \cc{L}_6(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} \ge 0 \\
        % semantic constraint of line 6 (else branch)
        \cc{L}_8(\cc{n}, \cc{abs}) &\gets \cc{L}_4(\cc{n}, \cc{abs}) \land \cc{n} < 0 \\
        % assign branch
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_6(\cc{n},\cc{abs}) \land \cc{abs'} = \cc{n} \\
        \cc{L}_9(\cc{n}, \cc{abs'}) &\gets \cc{L}_8(\cc{n},\cc{abs}) \land \cc{abs'} = -\cc{n} \\
        % semantic constraint of main procedure:
        \cc{main}(\cc{n},\cc{res}) &\gets \cc{L}_9(\cc{n}, \cc{abs'}) \land
        \cc{dbl}(\cc{x}, \cc{d})
        \land \cc{abs'} = \cc{x} \land \cc{res} = \cc{d} \\
        %
        False &\gets \cc{main}(\cc{n}, \cc{res}) \land \cc{res} < 0 
      \end{align}
    }
  \end{floatrow}
\end{figure}
% walk through code of running example:
%
\cc{dblAbs} is a procedure that doubles the absolute value of
its input and stores the result in \cc{res}.
% state the safety problem:
The program also asserts that \cc{res} is greater than or equal to $0$
before exiting.
% talk about solving Horn Clauses:
Verifying this assertion reduces to solving a recursion-free
CHC system over a set of uninterpreted predicates that represent
the control locations in \cc{dblAbs}.
%
In particular, one such system, $\mcchc$, is shown in
\autoref{fig:chc}.
%
%
While $\mcchc$ has been presented as the result
of a translation from \cc{dblAbs}, \sys is
purely a solver for CHC systems: it does not require access to the
concrete representation of a program, or for a given CHC system to be
the result of translation from a program at all.

@@

For example, consider $\cc{L}_9$. By the time \sys attempts to
synthesize an interpretation for $\cc{L}_9$ it will have solutions for
$\cc{L}_4$, $\cc{L}_6$, $\cc{L}_8$.
%
Possible interpretations of these predicates are shown in
\autoref{fig:ex-graph}.
%
The pre-formula is constructed from the bodies of clauses where
$\cc{L}_9$ is the head.
%
Each relational predicate, $P$,
is replaced by a corresponding boolean indicator variable, $\cc{b}_P$.
%
Each boolean indicator variable implies the solution for its
predicate,
encoded as the disjunction of the negation of the boolean
indicator variable and the solution.
%
In particular, the pre-formula for $\cc{L}_9$ is constructed from
clauses (5) and (6):
\begin{gather}
  ((\cc{b}_{\cc{L}_6} \land \cc{abs'} = \cc{n})
  \lor
  (\cc{b}_{\cc{L}_8} \land \cc{abs'} = -\cc{n}))
  \land
  (\neg \cc{b}_{\cc{L}_6} \lor \cc{n} \ge 0)
  \land
  (\neg \cc{b}_{\cc{L}_8} \lor \cc{n} < 0)
\end{gather}
%
The post-formula is constructed from clauses that transitively depend
on $\cc{L}_9$. Again, we replace relational predicates by
corresponding boolean indicators. However, we omit the boolean indicator for
$\cc{L}_9$. The post-formula is composed from clauses (1), (7), and
(8):
\begin{gather}
  (\neg \cc{b}_{\cc{dbl}} \lor \cc{d} = 2*\cc{x})
  \land
  (\neg \cc{b}_{\cc{main}} \lor
    (\cc{b}_{\cc{dbl}}
    \land \cc{abs'}= \cc{x}
    \land \cc{res}=\cc{d} ))
  \land
  (\cc{b}_{\cc{main}} \land \cc{res} < 0)
\end{gather}
%
Interpolating the pre and post formulas yields an intepretation
of $\cc{L}_9$: $\cc{abs'} \geq 0$.
%
The procedure for solving a CDD system is described in formal detail
in \autoref{sec:solve-cdd}.

@@

%% talk about expansion:
\begin{figure}[t]
  \centering
  \begin{floatrow}[2]
    \ffigbox[.45\textwidth] {%
      {\caption{$\mcchc$ as a directed hypergraph. Each relational
          predicate is depicted as a graph node while each clause is
          represented by a hyperedge. Each hyperedge is labelled by the
          fact in the corresponding CHC\@. Each node has a valid
          corresponding interpretation, written in braces.
      }\label{fig:ex-graph}}}
      {\input{fig/example}}
    \ffigbox[.45\textwidth]
      {\caption{The hierarchy of classes of recursion-free CHC
          systems. Body Disjoint and Linear systems are subsumed by
          CDD systems. Solving Directly Solvable CHC systems is in
          co-NP while solving general, recursion-free systems is
          co-NEXPTIME Complete.
      }\label{fig:hierarchy}}
      {\input{fig/hierarchy}}
  \end{floatrow}
\end{figure}
